### 2018-07-08 Chapters/Chapter1.tex
Heisenber model

Additionally, for $i=j=k=l$ the Pauli principle requires $\sigma' = \sigma$.

@@

\begin{equation}
\hat{H} = -t\sum_{\langle i,j \rangle, \sigma}(\hat{c}_{i \sigma}^\dagger \hat{c}_{j \sigma} + h.c.) + U \sum_{i=1}^M \hat{n}_{i\uparrow}\hat{n}_{i\downarrow}
\end{equation}

>>>>

Additionally, for $i=j=k=l$ the Pauli principle requires $\sigma' = \sigma$

@@

\begin{equation}
\label{Hubbard}
\hat{H} = -t\sum_{\langle i,j \rangle, \sigma}(\hat{c}_{i \sigma}^\dagger \hat{c}_{j \sigma} + h.c.) + U \sum_{i=1}^M \hat{n}_{i\uparrow}\hat{n}_{i\downarrow}
\end{equation}

\begin{section}{Heisenberg antiferromagnet from Hubbard model}

At half filling, in the $t << U$ limit we can apply perturbation theory to obtain the Heisenberg model from the Hubbard Hamiltonian \ref{Hubbard}. First notice that in the case $t = 0$ the ground state corresponds to having all sites single-occupied. There are two possible spin orientations for each sites, so there is a $2^M$ degeneracy. The ground state energy is $0$. The first exited states are obtained by moving one electron from one site to another, thus leaving one empty site and one double-occupied site. The energy of one of these states is $U$. Now, if we turn the kinetic term on as a perturbation, having two neighbors electrons with antiparallel spin will allow them to hop and reduce the kinetic energy, whereas while being parallel this hopping is not allowed and there is no energy reduction. Therefore, we can see that when turning the kinetic term on antiparallel alignment will be favored.

Let us write $\hat{H} = \hat{H_0} + \hat{H_t}$, $\hat{H_t} = -t\sum_{\langle i,j \rangle, \sigma}(\hat{c}_{i \sigma}^\dagger \hat{c}_{j \sigma} + h.c.)$. Let $\ket{\phi}$ and $\ket{\phi'}$ be state from the ground state, then we know $\bra{\phi'} \hat{H_0} \ket{\phi} = 0$ and in first order in $t$ $\bra{\phi'} \hat{H} \ket{\phi} = \bra{\phi'} \hat{H_t} \ket{\phi} = 0$ since $\hat{c}_{i \sigma}^\dagger \hat{c}_{j \sigma}$ moves an electron from site $j$ to $i$ thus leaving $i$ doubly occupied, therefore $\hat{H_t} \ket{\phi}$ has no superposition with $\phi'$. In second order we have:

\begin{equation}
\bra{\phi'} \hat{H} \ket{\phi} = \sum_s \frac{\bra{\phi'} \hat{H_t} \ket{s}\bra{s} \hat{H_t} \ket{\phi}}{E_0-E_s}
\end{equation}

The sum being over all excited states. Since $\hat{H_t}\ket{\phi}$ is a superposition of states with exactly one double occupied site and one empty site, we see that only excited states with one double occupied site will contribute. These states have energy $E_s = U$ therefore we have $\bra{\phi'} \hat{H} \ket{\phi} = -\frac{1}{U}\sum_s \bra{\phi'} \hat{H_t} \ket{s}\bra{s} \hat{H_t} \ket{\phi} = \bra{\phi'} -\frac{\hat{H_t^2}}{U} \ket{\phi}$. These means that within the subspace generated by the ground states of $\hat{H_0}$, $\hat{H}$ acts as an effective Hamiltonian $\hat{H}_{eff} = -\frac{\hat{H_t^2}}{U}$. Now,

\begin{align*}
	\bra{\phi'} -\frac{\hat{H_t^2}}{U} \ket{\phi} =
	\bra{\phi'} -\frac{t^2}{U}\sum_{\langle i,j \rangle, \sigma} \sum_{\langle i',j' \rangle, \sigma'} \hat{c}_{i \sigma}^\dagger 			\hat{c}_{j \sigma} \hat{c}_{i' \sigma'}^\dagger \hat{c}_{j' \sigma'} \ket{\phi} =
	\bra{\phi'} -\frac{t^2}{U}\sum_{\langle i,j \rangle, \sigma, \sigma'} \hat{c}_{i \sigma}^\dagger \hat{c}_{j \sigma} \hat{c}_{j \sigma'}^\dagger \hat{c}_{i \sigma'} \ket{\phi} &= \\
	\bra{\phi'} -\frac{t^2}{U}\sum_{\langle i,j \rangle, \sigma, \sigma'} \hat{c}_{i \sigma}^\dagger \hat{c}_{i \sigma'} \hat{c}_{j \sigma'}^\dagger \hat{c}_{j \sigma'} \ket{\phi} &
\end{align*}

Where we imposed $i=j'$ and $j=i'$ because that's the only way $\hat{H_t^2}\ket{\phi}$ has no double occupied sites and so the matrix element does not vanish. Let us now introduce the on-site spin operators:

\begin{equation}
\boldsymbol{S}_i = \frac{1}{2} \sum_{\sigma, \sigma'} \hat{c}_{i \sigma}^\dagger \boldsymbol{\sigma}_{\sigma, \sigma'} \hat{c}_{i \sigma'}
\end{equation}

Which satisfy:

\begin{align*}
\hat{c}_{i \sigma}^\dagger \hat{c}_{i \sigma'} &= \delta_{\sigma \sigma'} \frac{1}{2} (n_{i \uparrow} + n_{i \downarrow}) + \boldsymbol{S}_i\boldsymbol{\sigma}_{\sigma', \sigma} \\ 
\hat{c}_{i \sigma} \hat{c}_{i \sigma'}^\dagger &= \delta_{\sigma \sigma'} \frac{1}{2} (2 - n_{i \uparrow} - n_{i \downarrow}) - \boldsymbol{S}_i\boldsymbol{\sigma}_{\sigma, \sigma'} 
\end{align*}

Introducing these relations into $H_{eff}$ and using $n_{i \uparrow} + n_{i \downarrow} = 1$ we obtain:

\begin{align*}
H_{eff} = -\frac{t^2}{U} \sum_{\langle i,j \rangle, \sigma \sigma'} \left(\frac{1}{2}\delta_{\sigma \sigma'} + \boldsymbol{S}_i				\boldsymbol{\sigma}_{\sigma' \sigma}\right)\left(\frac{1}{2}\delta_{\sigma \sigma'}-\boldsymbol{S}_j\boldsymbol{\sigma}				_{\sigma \sigma'}\right) = \\
-\frac{t^2}{U}\sum_{\langle i,j \rangle, \sigma \sigma'} \frac{1}{4}\delta_{\sigma \sigma'} - \sum_{ab} S_i^aS_j^b 							\boldsymbol{\sigma}_{\sigma'\sigma}^a \boldsymbol{\sigma}_{\sigma \sigma'}^b = 
-\frac{t^2}{U}\sum_{\langle i,j \rangle} \frac{1}{2} - \sum_{ab} S_i^aS_j^b	2\delta_{ab} &= \\
-\frac{t^2}{U}\sum_{\langle i,j \rangle} \frac{1}{2} - 2\boldsymbol{S}_i\boldsymbol{S}_j
\end{align*}

The constant term can be neglected leaving:

\begin{equation}
H_{eff} = \frac{2t^2}{U} \sum_{\langle i,j \rangle} \boldsymbol{S}_i\boldsymbol{S}_j
\end{equation}

Which is the Heisenberg model for antiferromagnets.
